\propertitle

### 2019-06-13 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-13T14:42:35+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

In our case, to apply the method to the total activity of the full
population we need to fix some of its expectation values, for example its
moments. But we don't have any measured averages for the full population to
equate our expectations to. Here enters point~\ref{item:sample_step}: the
probability calculus gives an exact relation between the first $m$ moments
for the full population and the first $m$ for the sample, so that the ones
determine the others and vice versa at every time bin; this is a classical
result of sampling theory \citep[for
example][\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
also][]{whitworth1897}. Combining this result with the maximum-entropy
prescription \enquote{expectations${}={}$measured averages} for the sample,
we have that the latter averages determine the moments for the full
population:
\begin{equation}
  \label{eq:chain_prescription_moments}
  \text{measured averages} \longrightarrow
  \text{sample moments} \longrightarrow
  \text{full-population moments}
\end{equation}

This step is even more straightforward if instead of power moments we use
\emph{normalized factorial moments}. The $m$th normalized factorial moment
is defined as the expectation
\begin{equation}
  \E\condt[\Bigg]{
    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
\end{equation}
where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
activity of a population of $n$ neurons. The $m$th factorial moment can be
interpreted as the expectation of the number of distinct $m$-tuples of
simultaneously active neurons, normalized by the maximum possible number of
distinct $m$-tuples.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-13T15:02:32+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

In our case, to apply the method to the total activity of the full
population we need to fix some of its expectation values, for example its
moments. But we don't have any measured averages for the full population to
equate our expectations to. Here enters point~\ref{item:sample_step}: the
probability calculus gives an exact relation between the first $m$ moments
for the full population and the first $m$ for the sample; the ones
determine the others and vice versa at every time bin. This is a classical
result of sampling theory \citep[for
example][\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
also][]{whitworth1897}. Combining this result with the maximum-entropy
prescription \enquote{expectations${}={}$measured averages} for the sample,
we have that the measured averages for the sample determine the moments for
the full population:
\begin{equation*}
  \label{eq:chain_prescription_moments}
  \mathrlap{\overbracket{\phantom{\text{measured averages} \rightarrow
        \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
  \text{measured averages} \rightarrow
        \underbracket{\text{sample moments}
  \rightarrow
    \text{full-population moments}}_{\textit{sampling theory}}
\end{equation*}

This step is even more straightforward if instead of power moments we use
\emph{normalized factorial moments}. The $m$th normalized factorial moment
is defined as the expectation
\begin{equation}
  \E\condt[\Bigg]{
    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
\end{equation}
where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
activity of a population of $n$ neurons. This moment can be interpreted as
the expectation of the number of distinct $m$-tuples of simultaneously
active neurons, normalized by the maximum possible number of distinct
$m$-tuples.
